### 2017-03-23 APPENDIX.tex
Added appendix page

\documentclass{article}
\usepackage{fullpage} % make the margins a reasonable size
\usepackage{amsmath}  % some handy math symbols
\usepackage{amssymb}  % more handy math symbols
\usepackage{amsthm}   % this is necessary to create theorems/lemmas/etc.
\usepackage{mathtools}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{pdfpages}
\usepackage{multicol}
\usepackage{ulem}

\pagenumbering{gobble}

\begin{document}
\title{APPENDIX}
\date{}
\maketitle
\end{document}